---
title: "SDS 315--Homework 9"
author: "Santhoshi Penmetsa--sp48982"
date: "2024-04-16"
output:
  pdf_document: default
  html_document: default
---

**[GitHub Link](https://github.com/santhoshipen/SDS315)**

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

```{r}
# Loading the libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(magrittr)
library(MatchIt)
```

```{r}
# Loading the datasets
turnout <- read.csv("turnout.csv")
solder <- read.csv("solder.csv")
```

# **Problem 1: Get out the vote**

## **Part A: How much more likely are GOTV call recipients to have voted in 1998?**

```{r}
# The proportion of those recieving a GOTV call who voted in 1998 
turnout_call <- filter(turnout, GOTV_call == 1) 
voted_1998 <- filter(turnout_call, voted1998 == 1) 
prop_call <- nrow(voted_1998)/nrow(turnout_call)
 
# The sample proportion of those not recieving a GOTV call who voted in 1998 
turnout_no_call <- filter(turnout, GOTV_call == 0)
voted_1998_2 <- filter(turnout_no_call, voted1998 == 1)
prop_no_call <- nrow(voted_1998_2)/nrow(turnout_no_call)

# A large sample 95% confidence interval 
diff_prop = prop_call - prop_no_call

prop1 = (prop_call* (1-prop_call))/nrow(turnout_call)
prop2 = (prop_no_call*(1 - prop_no_call))/nrow(turnout_no_call)

diff_prop_se <- sqrt(prop1 + prop2)
confint_upper <- diff_prop + (1.96 * diff_prop_se)
confint_lower <- diff_prop - (1.96 * diff_prop_se)
```

In the preliminary analysis, we were able to find out that the proportion of those recieving a GOTV call who voted in 1998 was **`r round(prop_call, 4) `**. The sample proportion of those not receiving a GOTV call who voted in 1998 was **`r round(prop_no_call, 4)`**. When the large-sample 95% confidence interval was calculated for the difference in these two proportions, the lower limit of the confidence interval was **0.1411** and the upper limit of the confidence interval was **0.2659**. 

## **Part B: Confounding Variables**

In this scenario, there are numerous confounding variables: **voted1996, AGE, and MAJORPTY**. These particular variables might be contributing to the likelihood of receiving a GOTV call and the likelihood of voting in 1998. We need to demonstrate that these variable are in fact associated with both receiving a GOTV call and voting in 1998. 

**voted1996 Variable**

```{r}
# Voted in 1996 and Voted in 1998
table1 <- xtabs(~voted1996 + voted1998, data = turnout) %>%
  prop.table(margin = 2) %>%
  round(4)

table2 <- xtabs(~voted1996 + GOTV_call, data = turnout) %>%
  prop.table(margin = 2) %>% 
  round(4)

data_frame1 <- as.data.frame.matrix(table1)
data_frame2 <- as.data.frame.matrix(table2)
data_frame1 <- cbind(c("Did Not Vote in 1996", "Did Vote in 1996"), data_frame1)
data_frame2 <- cbind(c("Did Not Vote in 1996", "Did Vote in 1996"), data_frame2)

# Creating the tables 
table1_styled <- kable(data_frame1, format = "latex", booktabs = TRUE, caption = "Proportions of voted1996 and voted1998", col.names = c("", "Did Not Vote in 1998", "Did Vote in 1998"), row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped", "hold_position"), latex_options = c("striped", "hold_position"))
table1_styled

table2_styled <- kable(data_frame2, format = "latex", booktabs = TRUE, caption = "Proportions of voted1996 and GOTV Call", col.names = c("", "Did Not Get a GOTV Call", "Did Get a GOTV Call"), row.names = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hold_position"), latex_options = c("striped", "hold_position"))
table2_styled

```

From Table 1, the proportion of those who voted in 1996 and in 1998 is **0.7624**, whereas as those who did not vote in 1996 but voted in 1998 is **0.2376**. This suggests that voting in 1996 does influence your voting behavior in 1998. From Table 2, the proportion of those voted in 1996 and recieved a call is **0.7126**, whereas the proportion of those who did not vote in 1996 and had not recieve a call is **0.4692**. This suggests that voting in 1996 does influence whether or not you recieve a GOTV call. Therefore, voted1996 makes it so that they are more likely to recieve a call and also more likely to vote, making it a confounder. 

**AGE variable**

```{r}
# Calculating the means of the age 
voted_in_1998 <- filter(turnout, voted1998 == 1)
avg_1998 <- mean(voted_in_1998$AGE)

no_voted_in_1998 <- filter(turnout, voted1998 == 0)
no_vote_avg <- mean(no_voted_in_1998$AGE)

call <- filter(turnout, GOTV_call == 1)
call_age <- mean(call$AGE)

no_call <- filter(turnout, GOTV_call == 0)
no_call_age <- mean(no_call$AGE)

```

Those who voted in the 1998 elections tend to be older. The average age among those who voted in 1998 is **`r round(avg_1998,3)`** and the average age among those who did not vote in 1998 is **`r round(no_vote_avg,3)`**. Since the average age for those voted in 1998 is higher than those who did not vote, this suggests that those are older are more likely to vote. In addition, the average age among those recieved the GOTV call is **`r round(call_age,3)`** and the average age among those who did not is **`r round(no_call_age,3)`**. Since the average age for those who recieved a call is higher than those who didn't, this suggests that older people are more likely to recieve GOTV calls. Thereofore, AGE variable makes it so that older people are more likely to vote and recieve a call, making it a confounder. 

**MAJORPTY variable**

```{r}
# Creating tables for the MAJORPTY variable
majorpty_table1 <- xtabs(~MAJORPTY + voted1998, data = turnout) %>%
  prop.table(margin = 2) %>%
  round(4)
majorpty_table2 <- xtabs(~MAJORPTY + GOTV_call, data = turnout) %>%
  prop.table(margin = 2) %>% 
  round(4)

df1 <- as.data.frame.matrix(majorpty_table1)
df2 <- as.data.frame.matrix(majorpty_table2)
df1 <- cbind(c("Not Registered as Member in a Political Party", "Registered as Member in  a Political Party"), df1)
df2 <- cbind(c("Not Registered as Member in a Political Party", "Registered as Member in  a Political Party"), df2)

majorpty_table1_styled <- kable(df1, format = "latex", booktabs = TRUE, caption = "Proportions of MAJORPTY and voted1998", col.names = c("", "Did Not Vote in 1998", "Did Vote in 1998"), row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped", "hold_position"), latex_options = c("striped", "hold_position"))
majorpty_table1_styled

majorpty_table2_styled <- kable(df2, format = "latex", booktabs = TRUE, caption = "Proportions of MAJORPTY and GOTV Call", col.names = c("", "Did Not Get a GOTV Call", "Did Get a GOTV Call"), row.names = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hold_position"), latex_options = c("striped", "hold_position"))
majorpty_table2_styled
```

From Table 3, the proportion of those who registered as a member of one of the major political parties and voted in 1998 is **0.8019**, whereas as those weren't registered and voted in 1998 is **0.1981**. This suggests that being registered as a member of a political party influences your voting behavior in 1998. From Table 4, the proportion of those registered as a member of a political party and recieved a GOTV call is **0.7448**, whereas those weren't registered and did not recieve a call is **0.2552**. This suggests that being registered as a member of a political party most likely increases your chances of recieving a GOTV call. Therefore, MAJORPTY variable makes it so that they are more likely to recieve a call and also more likely to vote, making it a confounder. 

## **Part C: Matching**

```{r}
# Making the matched dataset
turnout_matched =  matchit(GOTV_call ~ factor(voted1996) + AGE + factor(MAJORPTY), data = turnout, ratio = 5) %>%
  match.data
```

**Contingency Table 1: Voted in 1996 vs. GOTV Call**
```{r}
# Proving voted1996 is not a confounder anymore
# GOTV: treatment, voted1998: predictor 
xtabs(~ voted1996 + GOTV_call, data = turnout_matched) %>%
                      prop.table(margin=2)
```

**Contingency Table 2: Age vs. GOTV Call**
```{r}
# Proving age is not a confounder anymore 
mean_call_age <- filter(turnout_matched, GOTV_call == 1)
age_call <- mean(mean_call_age$AGE)

mean_no_call_age <- filter(turnout_matched, GOTV_call == 0)
age_no_call <- mean(mean_no_call_age$AGE)

mean_table <- data.frame(Category = c("Mean Age with GOTV Call", "Mean Age without GOTV Call"),
                         Mean_Age = c(age_call, age_no_call))
mean_table
```

**Contingency Table 3: Registered to Major Party vs. GOTV Call**
```{r}
# Proving MAJORPTY is not a confounder anymore 
 xtabs(~ MAJORPTY + GOTV_call, data = turnout_matched) %>% 
  prop.table(margin = 2)
```

```{r}
# The proportion of those recieving a GOTV call who voted in 1998 
call_matched <- turnout_matched %>% 
  filter(GOTV_call == 1)
call_vote_matched <- call_matched %>%
  filter(voted1998 == 1)
call_prop_matched <- nrow(call_vote_matched)/ nrow(call_matched)
  

# The sample proportion of those not recieving a GOTV call who voted in 1998 
no_call_matched <- turnout_matched %>% 
  filter(GOTV_call == 0)
no_call_vote_matched <- no_call_matched %>%
  filter(voted1998 == 1)
no_call_prop_matched <- nrow(no_call_vote_matched)/ nrow(no_call_matched)

# A large sample 95% confidence interval 
prop_diff = call_prop_matched - no_call_prop_matched

prop_1 = (call_prop_matched* (1-call_prop_matched))/nrow(call_matched)
prop_2 = (no_call_prop_matched*(1 - no_call_prop_matched))/nrow(no_call_matched)

diff_prop_se <- sqrt(prop_1 + prop_2)
confint_lower <- prop_diff - (1.96 * diff_prop_se)
confint_upper <- prop_diff + (1.96 * diff_prop_se)
```

In the tables above, 0 represents No and 1 represents Yes. As observed in the tables, the proportions of if they voted in 1996, age, and major party registration seem to balanced with if they got a GOTV call. Therefore, this shows that the "matched" data is in fact balanced and accounts for the confounding variables previously explored.

For the matched dataset, the proportion of individuals who received a GOTV call and voted in 1998 is approximately **`r round(call_prop_matched, 4)`** The sample proportion of individuals who did not receive a GOTV call and voted in 1998 is approximately **`r round(no_call_prop_matched, 4)`**.When the large-sample 95% confidence interval was calculated for the difference in these two proportions, the lower limit of the confidence interval was **0.0056** and the upper limit of the confidence interval was **0.1418**. Since a high proportion of those who received a GOTV call seemed to have voted in the 1998 election, it can be concluded that GOTV call appears to have a positive impact on the likelihood of voting in the 1998 election. 

# **Problem 2: Manufacturing flaws in circuit boards**

## **Part A**

```{r}
# Making the opening typs vs. number of skips plot
ggplot(solder) + 
  geom_boxplot(aes(x = Opening, y = skips, fill = factor(Opening)), position = "dodge") +
  facet_grid(. ~ Opening, scales = "free_x", space = "free_x", labeller = labeller(Opening = c('L' = '', 'M' = '', 'S' = ''))) + 
  labs(title = 'Opening Type vs. Number of Skips', x = 'Opening Type', y = 'Number of Skips', fill = 'Opening Type') + 
  scale_fill_manual(name = 'Opening Type', labels = c('Large', 'Medium', 'Small'), values = c('hotpink', 'skyblue', 'violet')) +
  scale_x_discrete(labels = c('', '', '')) +  # Set x-axis labels
  theme_minimal() + 
  theme(text = element_text(size = 13))
```

The graph above shows different opening types and the range of the number of skips on the circuit board.

```{r}
# Making the solder type vs. number of skips plot
ggplot(solder) + 
  geom_boxplot(aes(x = Solder, y = skips, fill = factor(Solder)), position = "dodge") +
  facet_grid(. ~ Solder, scales = "free_x", space = "free_x", labeller = labeller(Solder = c('Thick' = '', 'Thin' = ''))) + 
  labs(title = 'Solder Type vs. Number of Skips', x = 'Solder Type', y = 'Number of Skips', fill = 'Opening Type') + 
  scale_fill_manual(name = 'Solder Type', labels = c('Thick', 'Thin'), values = c('purple', 'pink')) +
  scale_x_discrete(labels = c('', '')) +  # Set x-axis labels
  theme_minimal() + 
  theme(text = element_text(size = 13))
```

The graph above shows different solder types and the range of the number of skips on the circuit board.

## **Part B**

```{r}
# Making the regression model 
rm_skips <- lm(skips ~ Opening:Solder + Solder + Opening, data = solder) 
rm_skips_coef <- coef(rm_skips) 

rm_skips_coef <- as.data.frame(rm_skips_coef)

row.names(rm_skips_coef) <- c("Thick Solder & Large Opening", "Thin Solder", "Medium Opening", "Small Opening", "Medium Opening + Thin Solder", "Small Opening + Thick Solder")

# Making the tables 
kable(rm_skips_coef, booktabs = TRUE, col.names = c("Coefficient Value")) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

rm_skips <- confint(rm_skips, level = 0.95)
row.names(rm_skips) <- c("Thick Solder & Large Opening", "Thin Solder", "Medium Opening", "Small Opening", "Medium Opening + Thin Solder", "Small Opening + Thick Solder")
kable(rm_skips, booktabs = T) %>% 
  kable_styling(latex_options = c("striped", "hold_position"))
```


## **Part C**

1. Thick Solder & Large Opening: The baseline expected change in the number of skips for circuit boards with thick solder and large opening is approximately 0.39 while holding the variables constant. 

2. Thin Solder: The expected change in the number of skips for circuit boards with thin solder in isolation is approximately 2.28 while holding the variables constant.

3. Medium Opening: The expected change in the number of skips for circuit boards with a medium-sized opening in isolation is approximately 2.41 while holding the variables constant.

4. Small Opening: The expected change in the number of skips for circuit boards with a small opening in isolation is approximately 5.13 while holding the variables constant.

5. Medium Opening + Thin Solder: Tthe expected change in the number of skips for circuit boards with a medium-sized opening and thin solder is approximately -0.74 while holding the variables constant.

6. Small Opening + Thick Solder: The expected change in the number of skips for circuit boards with a small opening and thick solder is approximately 9.65 while holding the variables constant. 

## **Part D**

Based on the analysis, my recommendation to AT&T would be to consider a medium Opening with a thin Solder. Since the coefficient -0.74 has the lowest estimated number of skips among all combinations, this would help reach the goal of minimizing the number of skips in the manufacturing process.

