---
title: "SDS 315--Homework 3"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2024-01-30"
---

**[GitHub Link](https://github.com/santhoshipen/SDS315)**

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(fig.height=5, fig.width=9, warning = FALSE, message = FALSE, echo = FALSE)
```

```{r}
#Loading the libraries 
library(ggplot2)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(RColorBrewer)
library(kableExtra)

#Loading the datasets
creatinine <- read.csv("creatinine.csv")
market_model <- read.csv("marketmodel.csv")
covid <- read.csv("covid.csv")
milk <- read.csv("milk.csv")
```

# **PROBLEM 1**

**Part A**
```{r}
#Creating a linear regression model 
model_creatine = lm(creatclear ~ age, data = creatinine)

#Printing out the coefficients of the model 
coef = round(coef(model_creatine),3)

#Calculating the creatinine level for the specific age 55 
calc = 147.813 - 0.620*55
calc
```

To estimate the creatinine clearance rate for a 55-year old, I used the linear regression model fitted to the data. 
The slope is approximately -0.62 and the y-intercept is 147.813. Then, the expected creatinine clearance rate was obtained by plugging in the value 55. 

**Part B** 
```{r}
#Printing out the slope of the linear regression model 
slope_age <- round(coef(model_creatine)[2],3)
slope_age
```

To determine how creatinine clearance rate changes with age, I examined the linear regression model fitted to the data, focusing on the coefficient associated with the 'age' variable. The coefficient value (-0.62) represents the projected change in creatinine clearance rate per one-unit increase in age, resulting in an average decrease of 0.62 mL/minute per year.
 
**Part C**
```{r}
#Calculating the creatinine level for the specific age 40
forty_year_old = 135 - (147.8129 - 0.6198*40)

#Calculating the creatinine level for the specific age 60
sixty_year_old = 112 - (147.8129 - 0.6198*60)
```
The forty year old's creatinine clearance rate is **`r round(forty_year_old,3)`** ml/min, and the sixty year old's creatinine clearance rate is **`r round(sixty_year_old,3)`** ml/min. 


The 40-year old's creatinine clearance rate is healthier. Using the linear regression model fitted to the data, the creatinine levels for their specific ages were calculated. The equation used to calculate was 147.8129 - 0.6198x, plugging in their respective ages and then subracting it from their actual creatinine clearance rate. This demonstrated that both of them surpassed what was predicted by the linear regression model. However, because the 40-year-old has a higher predicted value, they are regarded as healthier. 

# **PROBLEM 2**

```{r}
#Using a for loop to create the regression table 
reg_table <- tibble()

stock_variables <- c("AAPL", "GOOG", "MRK", "JNJ", "WMT", "TGT")

for (stock in stock_variables) {
  model_stock <- lm(get(stock) ~ SPY, data = market_model)
  
  results <- tibble(
    Symbol = stock,
    Intercept = format(coef(model_stock)[1],digits = 3, scientific = FALSE),
    Slope = format(coef(model_stock)[2], digits = 3, scientific = FALSE),
    Rsquared = format(summary(model_stock)$r.squared, digits = 3, scientific = FALSE),
  )
  
  reg_table <- rbind(reg_table, results)
}

#Using the kable_styling to make the table professional 
kbl(reg_table, booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position"), position = "left", full_width = FALSE)

```

Beta is defined as the percentage change in an asset's return resulting from a one percent change in the market portfolio. It is essentially a measure of the riskiness of a stock relative to the wider market. When considering the risk of a company's stock, there are two classifications: unsystematic and systematic. Unsystematic risk is firm-specific and may be diversified away; systematic risk is market-wide and cannot be diversified away. And, beta is a measure of this systemic risk. If a stock's beta is greater than one, it has more systematic risk than the average. If a stock's beta is smaller than one, the systematic risk is lower. If a stock's beta is zero, there is no systematic risk because it does not respond to marketwide movements. A negative beta means that the firm has a negative risk premium. In the context of a regression model, beta can usually be found through a regression analysis on historical price data for the stock and its index of reference. The regression model evaluates the relationship between stock returns and market returns. The slope of the regression line shows the beta coefficient, which reflects the stock's susceptibility to market movements. In essence, beta informs investors how much a stock's price is likely to change in a reaction to market changes, enabling them to assess risk and make better investment choices.

The table above displays the regression of the returns for each of the 6 stocks individually on the return of S&P 500. It consists of beta (slope), alpha(intercept), and Rsquared for each of the 6 regressions. 

WMT has the lowest systematic risk because it has the smallest slope/beta (0.519), which indicates lower susceptibility to market movements. AAPL has the highest systematic risk because it has the highest slope/beta (1.07) and it's over 1, which indicates higher susceptibility to market movements. 


# **PROBLEM 3**

```{r}
#Filtering the dataset based on the country
covid_italy = filter(covid, country == "Italy")
covid_spain = filter(covid, country == "Spain")

#Using the linear regression models 
lm_italy = lm(log(deaths) ~ days_since_first_death, data=covid_italy)
italy_coef = coef(lm_italy)
lm_spain = lm(log(deaths) ~ days_since_first_death, data=covid_spain)
spain_coef = coef(lm_spain)

#Finding out the growth rates for each country
growth_rate_italy = coef(lm_italy)[2]
growth_rate_spain = coef(lm_spain)[2]

#Finding out the doubling time for each country
doubling_time_italy = round(70/(growth_rate_italy * 100),)
doubling_time_spain = round(70/(growth_rate_spain * 100),)

```

For Italy, the estimated growth rate is **`r round(growth_rate_italy,3)`** and the doubling time is **`r doubling_time_italy`** days. For Spain, the estimated growth rate is **`r round(growth_rate_spain,3)`** and the doubling time is **`r doubling_time_spain `** days. 


```{r}
#Plotting the line graph 
ggplot(covid) + geom_line(aes(x = days_since_first_death, y = deaths, color = country)) + labs(title = "Reported Daily  Deaths Over Time", x = "Days Since First Death", y = "Number of Deaths")
```

The line graph above displays the reported daily COVID related deaths over time in two countries: Spain and Italy. The number of deaths in Spain seems to have increased faster than in Italy.

# **PROBLEM 4**

```{r}
#Using the log-log scale to find the coefficients 
lm_milk = lm(log(sales) ~ log(price), data=milk)
milk_coef = coef(lm_milk)

#Finding out the elasticity of the milk 
milk_elasticity = coef(lm_milk)[2]
```

Th estimated price elasticity of demand for milk is **`r round(milk_elasticity,3)`**. To determine this, I ran the regression on the log-log scale. The outputted coefficient of log(price) was the estimated elasticity of demand for milk, which indicates how milk purchases are influenced by changes in price.
