---
title: "SDS 315--HOMEWORK 2"
author: "Santhoshi Penmetsa--sp48982"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2024-01-18"
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(fig.height=5, fig.width=9, warning = FALSE, message = FALSE, echo = FALSE)
```

```{r}
#Loading the libraries 
library(ggplot2)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(RColorBrewer)

#Loading the datasets
profs <- read.csv("profs.csv")
bikes <- read.csv("bikeshare.csv")
metro <- read.csv("capmetro_UT.csv")
billboard <- read.csv("billboard.csv")
```

# **PROBLEM 1: BEAUTY, OR NOT, IN THE CLASSROOM**

# **Part A**
```{r}
##Creating a barplot for the distribution of course evaluations
ggplot(profs) + geom_histogram(aes(x=eval), bins = 30, col = 'black', fill= '#6d57ca') + labs(title= 'Distribution of Course Evaluations', x= 'Course Evaluations', y= 'Frequency')
```

The above histogram depicts the distribution of course evaluation scores. The distribution of course evaluation scores appears to be slightly skewed to the left. The median evaluation score is `r median(profs$eval)`.The range of the score distribution is `r max(profs$eval) - min(profs$eval)`.The key takeaway is that students tend to evaluate professors with generally high scores. 

# **Part B**
```{r echo = FALSE}
#Creating a boxplot that shows the relationship between speaking status and evaluations scores
ggplot(profs) + geom_boxplot(aes(x=native,y = eval), col = 'black', fill = '#f19cbb') + labs(title = 'Course Evaluations Scores by Native Speaking Status', x = 'Native English Speaking Status (Yes or No)', y = 'Course Evaluation Scores')
```

The side-by-side boxplots above show the distribution of course evaluation scores based upon whether the professor is a native English speaker. Professors who spoke English natively had a higher median score than those who did not. Professors who were not native English speakers had no outliers, but native speakers had several low outliers. The key takeaway here is that professors who are native English speakers have varied course evaluation scores as the boxplot is more spread out. 

# **Part C**
```{r echo = FALSE, warning = FALSE, message = FALSE}
#Creating a faceted bar plot for distribution of course evaluation and gender
ggplot(profs) + geom_histogram(aes(x = eval), col = 'black', fill = '#5e9655') + facet_wrap(~gender) + labs(title = "Distribution of Course Evaluations by Professor's Gender", x = 'Course Evaluations', y = "Professor's Gender")
```

The faceted histogram shows a comparison of the distribution of course evaluation scores dependent on the gender of the professor. The distribution of scores for male professors appears to be more skewed to the left than for female professors. The key takeaway is that male professors had a higher concentration of 4 and 4.5 score ratings than the female professors. 

# **Part D**
```{r echo = FALSE}
#Creating a scatterplot to show the relationship between evaluation scores and attractiveness
ggplot(profs) + geom_point(aes(x = beauty, y = eval), col = '#43c0f6') + labs(title = 'Course Evaluations by Attractiveness', x = 'Physical Attractiveness', y = 'Course Evaluations')
```

The scatter plot above displays the relationship between physical attractiveness and course evaluation scores. `r round(cor(profs$beauty, profs$eval),2)`. This indicates there is no relationship. The key takeaway here is that the physical attractiveness of the professor does not influence their score rating.

# **PROBLEM 2: BIKE SHARING**

# **Part A**
```{r echo = FALSE}
#Computes the average bike rentals by grouping the 'hr' variable
avg_rentals <- bikes %>% group_by(hr) %>% summarise(avg_total = mean(total))

#Creates a line graph of the average hourly bike rentals 
ggplot(avg_rentals) + geom_line(aes(x = hr, y = avg_total), col = '#8679b4') + labs(title = 'Average Hourly Bike Rentals', x = 'Hour of the Day', y = 'Average Bike Rentals')
```

This line graph shows the average rentals across all hours of day. The average bike rentals peaked between the 15th and 20th hours of the day, with the second biggest peak being between the 5th and 10th hours. The lowest average bike rentals occurred within the first 5 hours of the day. The key takeaway here is that people rent bikes more frequently later in the day. 

# **Part B**
```{r echo = FALSE}
#Merges the dataset and creates a new variable to facet by workday
avg_rentals_merged <- merge(avg_rentals, bikes)
avg_rentals_merged$type <- ifelse(avg_rentals_merged$workingday == 1, "Workday", "No Workday")

#Creates a line graph by grouping the 'type' and 'hr' variables 
avg_rentals_merged %>% group_by(type, hr) %>% summarise(avg_total = mean(total)) %>% ggplot() + geom_line(aes(x = hr, y = avg_total, color = type)) + labs(title = "Average Hourly Bike Rentals by Working Day", x = "Hour of the Day", y = "Average Total Bike Rentals", color = 'Day Type') + scale_color_manual(values = c('Workday' = '#d63232', 'No Workday' = '#003255')) + theme_minimal() + facet_wrap(~type, scales = 'free_y')

```

The faceted line graph above displays average rentals for all hours of the day based on whether it is a workday. The number of average bike rentals throughout a workday peaks at two separate times, with a decline in between. During a non-working day, the number of average bike rentals has only one peak. The key takeaway is that individuals are more likely to rent out bikes consistently throughout the day during non-workday than on workdays. 

# **Part C**
```{r echo = FALSE} 
#Filtering and grouping by the 'weathersit' and 'workingday' variables 
bikes9am<- filter(bikes, hr == 9)
avg_bikes9am <- bikes9am %>% group_by(weathersit, workingday) %>% summarise(avg_total = mean(total))

#Creating a bar plot that shows the average ridership by weather situation at 9 am by day type 
ggplot(avg_bikes9am) + geom_bar(aes(x = weathersit, y = avg_total, fill = factor(workingday)), stat = 'identity') + facet_wrap(~workingday, labeller = labeller(workingday = c('0' = 'No', '1' = 'Yes'))) + labs(title = 'Average Ridership by Weather Situation at 9 am by Weekday', x = 'Weather Situation', y = 'Total Bike Rentals', fill = 'Working Day') + scale_fill_manual(name = 'Working Day', labels = c('No', 'Yes'), values = c('#b58c88', '#707d7f')) + theme(legend.position = "none") + theme_minimal()
```

The faceted barplot above shows average bike rentals by weather situation at 9 a.m. based on whether it is a workday or not. During all weather conditions, there was a higher average ridership on working days compared to non-working days. The average ridership on a working day was comparable on clear, cloudy days and misty, broken clouds days. The key takeaway is that more average bike rentals occur on working day and the weather situation does affect the ridership.  

# **PROBLEM 3: CAP METRO UT RIDERSHIP**

```{r echo = FALSE }
#Reordering the days, months and week 
metro = mutate(metro, day_of_week = factor(day_of_week, levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")), month = factor(month,levels=c("Sep", "Oct","Nov")))
```

# **Plot A** 
```{r echo = FALSE}
#Grouping by the 'hour_of_day', 'month' and 'day_of_week'
metro_data <- metro %>% group_by(hour_of_day, month, day_of_week) %>% summarise(avg_boarding = mean(boarding))

#Creating a line graph of average hourly boardings by month and week of day 
ggplot(metro_data) + geom_line(aes(x = hour_of_day, y = avg_boarding, color = month)) + facet_wrap(~ day_of_week) +  labs(title = 'Average Hourly Boardings by Month and Week of Day',x = "Hour of the Day", y = "Average Boardings", color = "Month") + theme_minimal()
```

The line graphs above describe the average boardings faceted by hour of day, month and day of the week. The average boardings are higher on weekdays(Mon-Fri) than the weekend(Sat-Sun).The peak boarding time is the same on all weekdays during the 15th hour. The average boardings are low in September on Mondays due to the federal holiday, Labor Day. Since it accounts for 1/4 of the Mondays in September, it shows the a decrease. Similarly, the average boardings on Weds/Thurs/Fri in November look lower because of the Thanksgiving as people are on leave or traveling. 

# **Plot B**
```{r echo = FALSE}
#Creating a scatterplot that is faceted by hour of day and day type
ggplot(metro) + geom_point(aes(x = temperature, y = boarding, color =  weekend), size = 0.5) + facet_wrap(~hour_of_day) + labs(title =  "Hourly Boardings vs. Temperature by Day Type", x='Temperature', y = 'Boarding') 
```

The scatterplot above depicts the distribution of temperature and boardings according to the hour of day and whether it is a weekend or weekday. Even when the time of day and week were kept constant, there would be no positive or negative correlation found between temperature and boardings. The number of CapMetro riders appears to be unaffected by the temperature.  

# **PROBLEM 4: WRANGLING THE BILLBOARD TOP 100** 

# **Part A** 
```{r echo = FALSE}
#Arranging the 'weeks_on_chart' variable in descending order
billboard <- arrange(billboard, desc(weeks_on_chart))

#Ensuring that top 10 songs are all unique and deleting irrelevant variables 
top_10 <- billboard %>% distinct(song, .keep_all = TRUE) %>% head(n = 10)
billboard.table <- subset(top_10, select = -c(X,url,week_id,week_position,song_id,instance,previous_week_position, year, week, peak_position))

#Naming the variables and creating the table 
names(billboard.table) = c("song", "performer", "count")
kable_styling(kbl(billboard.table) ,full_width = FALSE, position = 'left')
```

The table above displays the ten most popular songs since 1958. The most popular song was "Radioactive" by Imagine Dragons, which spent 87 weeks on the list. There were two distinct pairs of songs that spent the same number of weeks on the chart. 

# **Part B**
```{r echo = FALSE}
#Filtering the data to exclude the years 1958 and 2021
billboard_filter <- filter(billboard, year %in% seq(1959,2020))

#Grouping by the 'year'and 'song_id' variables 
billboard_filter_songs <- billboard_filter %>% group_by(year, song_id) %>% summarise(count = n())
songs <- billboard_filter_songs %>% group_by(year) %>% summarise(unique.songs = n())

#Creating a line graph that shows the musical diversity
ggplot(songs) + geom_line(aes(x = year, y = unique.songs), col = '#a8a5ec') + labs(title = 'Musical Diversity of Billboard 100 Across the Years', x = 'Year', y = 'Number of Unique Songs') + theme_minimal()
```

The line graph above shows how musical diversity has changed over time. This graph excludes the years 1958 and 2021. The largest peak in the number of unique songs occurred between 1960 and 1980.  Since roughly 1970, the number of unique scores has declined, reaching a low in the early 2000s. Since the 2000s, there has been a gradual increase. 

## **Part C** 
```{r echo = FALSE}
#Filtering the data 
ten_week  <- filter(billboard, weeks_on_chart >= 10)

#Grouping the data and ensuring there are unique songs 
ten_week_songs <- ten_week %>% group_by(performer, song) %>% summarise(count = n_distinct(song)) %>% group_by(performer) %>% summarise(num_of_songs = sum(count))

#Filtering the data and creating a bar plot of ten week hits
ten_week_hits <- filter(ten_week_songs, num_of_songs >= 30)
ggplot(ten_week_hits) + geom_bar(aes(x = performer, y = num_of_songs), stat = "identity", col = 'black',fill = 	'#e68fac') + coord_flip() + labs(title = 'Number of Ten Week Hits for 19 Artists', x = 'Performer', y = 'Number of Songs')
```

The bar plot above depicts the amount of ten-week hits for 19 different artists. Each of these musicians has had at least 30 ten-week hits. Elton John had the most ten-week hits, with more than 50 songs. Madonna comes in second with close to 45 tracks. There are about 5 musicians with exactly 30 ten-week hits.
